---
title: "Rainfall Data Reporting for `r params$region`"
author: Barry Rowlingson
output: pdf_document
params:
  datafolder: NA
  region: NA
---

```{r note, eval=FALSE, echo=FALSE}
Notes:

    * This document should render without any extra packages other
       than knitr, rmarkdown and ggplot2.
    
    * This document requires two parameters: the data folder and the region name:

    render("reporting-report.Rmd", params=list(datafolder="./DATA/", region="Lancaster"))
    
```

```{r init, echo=FALSE}
## hide all code, attach packages
knitr::opts_chunk$set(echo=FALSE)
library(ggplot2)
```


```{r rrs}
## read the data from the two files in the given
## folder and return in a list
read_readings_stations <- function(datapath){
    
### get the readings
    readings = read.csv(file.path(datapath, "readings.csv"))
### change the month to a factor
    readings$month = factor(readings$month, levels=month.name)
### contruct an ISO8601 year-month-day column
    readings$ymd = paste0(readings$year,"-",readings$month,"-01")
### convert to a Date object
    readings$date = as.Date(readings$ymd, format="%Y-%B-%d")

### Get the stations
    stations = read.csv(file.path(datapath, "stations.csv"))
    
    return(
        list(
            readings=readings,
            stations=stations
        )
    )
}

### read from the data folder passed as parameter
data = read_readings_stations(params$datafolder)

### break out into two objects for convenience
readings = data$readings
stations = data$stations    
```


```{r aggregations}
### make some summary aggregation tables:

## mean rainfall given a date
date_mean = aggregate(rain~date, readings, mean, na.rm=TRUE)
## fix the months as factors
date_mean$month = factor(format(date_mean$date, "%B"), levels=month.name)
## extract the year from the date
date_mean$year = as.numeric(format(date_mean$date,"%Y"))

## mean rain for each of the 12 calendar months
month_mean = aggregate(rain~month, readings, mean)

## total rainfall in each year measured by each station
annual_station_total = aggregate(rain ~ stationID+year, readings, sum)
```

## Data Summary

The data consists of `r nrow(readings)` rainfall records from 
`r nrow(stations)` station locations. The data period is from `r min(readings$year)` to
`r max(readings$year)`.

## Rainfall Over Time

The following plot shows the mean monthly rainfall for all stations
reporting in a month.

```{r plotdate}
plot(rain~date, date_mean, ylab="inches",type="l")
# or ggplot(date_mean,aes(x=date, y=rain)) + geom_line()
```

This plot shows the same data as in the previous plot, but as
a grid with month on the Y axis and year on the X axis. Any seasonal
difference between a dry and wet season should be more apparent.

```{r plotmr, fig.height=2.4}
## use a scale of increasing blue for wetness
ggplot(date_mean, aes(x=year, y=month, fill=rain)) +
    geom_raster() +
    scale_fill_distiller(direction=1)
```

## Monthly Averages

This graph shows the average over all stations and years for each
calendar month.

```{r plotmonth}
## rotate the axis labels using `las=2`. Add the rainfall
## measurement using `text`
b = barplot(rain~month, month_mean,las=2,xlab="",axes=FALSE)
text(b, 0.2, as.character(round(month_mean$rain,2)),cex=0.8)
abline(h=0) # neatens it up
axis(2) # scale on the left
```

## Maxima and Minima

This table shows the five wettest months recorded at any station.

```{r highest}
## need names of the stations, so do a merge on the station ID
full_data = merge(readings, stations, by="stationID")

## sort the table by rain in decreasing order
sortread = full_data[order(full_data$rain, decreasing=TRUE),]
## take the top 5 and make a table with four columns in this
## order
top5 = sortread[1:5,]
kable(top5[,c("name","month","year","rain")], row.names=FALSE)
```

## Wettest and Dryest Years

The five wettest years as measured at a single station are shown
in the next table.

```{r wettestinayear}
## merge the annual totals with the station data to get names
full_annual = merge(annual_station_total, stations, by="stationID")
## sort by name and do a table of the top 5
sorted_annual = full_annual[order(full_annual$rain, decreasing=TRUE),]
kable(sorted_annual[1:5, c("name","year","rain")], row.names=FALSE)
```

The five driest years at any station are in the next table. Low
values may be due to missing month records at a station.

```{r tailannual}
## get the last 5 rows using `tail`, and reverse the order
## so the lowest appears first
lowest = tail(sorted_annual, 5)[,c("name","year","rain")][5:1,]
kable(lowest, row.names=FALSE)
```

## Monthly Extremes

```{r extremestat}
## I want to say how many years of data we have in the text,
## so I compute this:
nyears = diff(range(readings$year))+1
```

In the `r nyears` years of the study data, each year appears as the 
wettest year the following number of times:

```{r wettest}
rsort = date_mean[order(date_mean$rain, decreasing=TRUE),]
wettest_month = rsort[!duplicated(rsort$year),]
kable(table(month=wettest_month$month))
```

As a graph of proportions of the total number of years, the
data looks like this:

```{r wettable}
barplot(table(wettest_month$month)/nyears,las=2)
abline(h=0) # add a baseline
```
